Modeling Macroeconomic Phenomena \\with Multi-Agent Reinforcement Learning

### 2019-12-07 paper/paper.tex

After verifying the preliminary results, we simulated a more complex scenario and tracking more standard economic metrics. We included five firm agents and 25 human agents using REINFORCE with episodes of length 200. The interest rate for human agents was set to $0.2\%$ and the operational cost for the firms was set to $\$10$. The money of each human agent for the first episode is shown in Figure \ref{plt:init-ppl}, and the money of each firm agent for the first episode is shown in Figure \ref{plt:init-firms}. The initial policies for human agents tended to buy goods greedily, while the initial policies for firm agents tended to sell greedily.

>>>>

After verifying the preliminary results to evaluate the basic success of the RL algorithms, we began to analyze individual policies chosen by firms and people as well as the macroeconomic effects that resulted from multiple agents interacting in the environment under more complicated dynamics with government policies.

Under the standard dynamics, human agents tend to converge on fairly trivial dynamics, immediately spending earned money on goods. Since the acquisition of goods ultimately drives reward, there is no incentive to save money, no matter the discount factor used. A particularly sharp demand curve for goods can impose a soft limit on a person’s buying rate, but otherwise does not change the strategy of spending early and exorbitantly.

A side effect of this phenomenon is a pronounced market collapse, or recession. When a person runs out of money due to excessive spending, they are unable to buy any additional goods. The firms, in turn, note the reduced demand for goods and begin to produce fewer goods. Accordingly, they employ fewer people to produce goods. As a result, unemployment increases, and lesser-skilled people can no longer find work. Without a source of income, they spend their remaining money and drop out of the market.

We found such recessions to be inescapable — the system can almost never recover, and the eventual equilibrium is one in which a single skilled person makes and buys all the goods -- a particularly unrealistic outcome. Since firms always up-sell goods (that is, sell them at a profit), it always requires fewer people to produce a set of goods than to consume them. The difference between these populations is thus unemployed and eventually unable to buy goods at all, and the employment rate steadily drops. People who lose all their money rarely interact in the market any further -- a situation eerily analogous to death.

Interestingly, although this outcome is not optimal for most people and firms, it proved to be very stable, and most randomly-initialized systems converged to a recession. To rectify this behavior, we increased the interest rate, thereby incentives the saving of money, which can purchase more goods after the accumulation of interest. Under this policy, we find that people naturally develop \textit{retirement policies}, under which they spend the first half of their lives saving more than they spend, and the latter half spending more than they save. Although such policies maximize people’s utility, they do not appear to be easy to converge to. Although many people happily save money and continually spend the accrued interest, relatively few people learn to truly ``enjoy'' their retirement -- dramatically increasing spending near the end of their lives.

As an example, we included five firm agents and 25 human agents using REINFORCE with episodes of length 200. The interest rate for human agents was set to $0.2\%$ and the operational cost for the firms was set to $\$10$. The money of each human agent for the first episode is shown in Figure \ref{plt:init-ppl}, and the money of each firm agent for the first episode is shown in Figure \ref{plt:init-firms}. As noted above, the initial policies for human agents tended to buy goods greedily, while the initial policies for firm agents tended to sell greedily.
